---
title: "Movie Statistics"
author: "Victoria Wingo"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    css: styles.css
    toc: true
    toc_float: true
    collapse: true
params:
  year: 2019
---

```{r data, include = FALSE}
library(ggplot2)
library(ggrepel)
library(dplyr)
library(readr)
library(DataCombine)
library(tidyverse)
library(knitr)
Highest_Hollywood_Grossing_Movies <- read.csv("Highest Hollywood Grossing Movies.csv")
```

# Datasets

## Movies

The `movies` dataset serves as a corrected version of the original dataset `Highest_Hollywood_Grossing_Movies` taken from kaggle, covering Hollywood films through 2021. Missing data has been added manually and the variable `runtime` has been converted from the 00h 00m format to minutes.

```{r movies, include = FALSE}
movies <- Highest_Hollywood_Grossing_Movies %>% 
  mutate(
    world_sales_mil = world_sales_dollars / 1000000, 
    release_year = format(as.Date(release_date, format="%B %d, %Y"),"%Y")) %>%
  select(-domestic_sales_dollars, -international_sales_dollars, -world_sales_dollars) %>%
  data.frame()

#replace original runtime format with minutes
split_run <- str_split(movies$runtime, " ")
run_ind <- 1
for(t in split_run){
  str_split(t, " ")
  # either hours or minutes missing
  if (length(t) == 2){
    if (t[2] == "min"){
      hr <- 0
      min <- as.integer(t[1])
    }
    else{
    hr <- as.integer(t[1])
    min <- 0      
    }
  }
  else{
    hr <- as.integer(t[1])
    min <- as.integer(t[3])
  }
  time <- (hr * 60) + min
  movies$runtime[run_ind] <- time
  run_ind <- run_ind + 1
}
movies$runtime <- as.integer(movies$runtime)
```

## Average Sales by Runtimes

```{r by_runtime, echo = FALSE, message = FALSE, fig.cap = "Fig. 1.1 On average, a longer movie will see more sales than a shorter movie, but sales become less predictable with longer movies."}
by_runtime <- movies %>% 
  group_by(runtime) %>% 
  summarize(avgSales = mean(world_sales_mil))
ggplot(by_runtime, 
       aes(x = runtime,
           y = avgSales)) + 
  geom_point() + 
  geom_smooth() + 
  labs(x = "Runtime (in minutes)", 
       y = "Average Sales (in millions USD)") + 
  scale_y_continuous(breaks = seq(from = 0, 
                                  to = (max(by_runtime$avgSales) + 50), 
                                  by = 200)) + 
  scale_x_continuous(breaks = seq(from = 0, 
                                  to = (max(by_runtime$runtime) + 30), 
                                  by = 30))
```

## Average Sales by License

```{r sales_by_license, echo = FALSE, fig.cap = "Fig. 2.1 On average, PG-13 movies sees more sales, while PG and G movies see about the same amount of sales."}
sales_by_license <- movies %>% 
  group_by(license) %>% 
  summarize(avgSales = mean(world_sales_mil))
ggplot(sales_by_license, 
       aes(x = license, 
           y = avgSales, 
           fill = license)) + 
  geom_col() + 
  guides(fill = "none") + 
  labs(x = "License", 
       y = "Average Sales (in millions USD)")
```

## Average Runtimes by License

```{r by_license, echo = FALSE, fig.cap = "Fig. 3.1 On average, PG-13 and R rated movies have about the same runtime."}
ggplot(movies, 
       aes(x = license, 
           y = runtime, 
           fill = license), 
       alpha = 0.6) + 
  geom_boxplot() + 
  guides(fill = "none") + 
  labs(x = "License", 
       y = "Runtime (in minutes)")
```

## Average Sales by Release Year

```{r by_year, echo = FALSE, fig.cap = "Fig. 4.1 From its early years through the 80's, average sales were less predictable, with 2020-2021 seeing a downturn due to the COVID pandemic."}
by_year <- movies %>% 
  group_by(release_year) %>% 
  summarize(avgSales = mean(world_sales_mil))
ggplot(by_year, 
       aes(x = release_year,
           y = avgSales)) + 
  geom_point() + 
  labs(x = "Year", 
       y = "Average Sales (in millions)") + 
  geom_text(data = subset(by_year, avgSales > 500 | avgSales <200), 
            aes(label = release_year, 
                hjust = 1.15, 
                vjust = -.5), 
            size = 3, 
            check_overlap = TRUE) + 
  scale_x_discrete(drop = FALSE,
                   expand = expansion(add = 3), 
                   breaks = seq(from = 1930, 
                                to = 2025,
                                by = 10)) + 
  scale_y_continuous(expand = expansion(mult = c(.1, .15)), 
                     breaks = seq(from = 0,
                                  to = (max(by_year$avgSales) * 1.15),
                                  by = 50))

movies %>% 
  filter(release_year == 1977) %>% 
  select(title, 
         release_date, 
         world_sales_mil, 
         license, runtime) %>% 
  kable(caption = "Fig. 4.2 Movies from 1977")

movies %>% 
  filter(release_year == 1980) %>% 
  select(title, release_date, world_sales_mil, license, runtime) %>% 
  kable(caption="Fig. 4.3 Movies from 1980")

movies %>% 
  filter(release_year == 2015 & world_sales_mil > 500) %>% 
  select(title, 
         release_date, 
         world_sales_mil, 
         license, 
         runtime) %>% 
  arrange(world_sales_mil) %>% 
  kable(caption = "Fig. 4.4 Movies from 2015 with world sales above 500 million USD")
```

## Movies with the Most World Sales of Their Year
```{r most_sales, echo = FALSE, warning = FALSE, fig.cap = "Fig. 5.1 These movies had the most sales of their year, showing a strong showing for superhero movies in the past 10 years."}
most_sales <- movies %>% 
  group_by(release_year) %>% 
  filter(world_sales_mil == max(world_sales_mil)) %>% 
  select(title, 
         release_year, 
         distributor,
         world_sales_mil) %>%
  arrange(desc(world_sales_mil))
ggplot(most_sales,
       aes(x = release_year,
           y = world_sales_mil)) +
  geom_point() +
  geom_text_repel(aes(label = title),
            size = 3,
            segment.color = "grey50",
            max.overlaps = 5) +
  scale_x_discrete(name = "Year",
                   breaks = seq(from = 1930,
                                to = 2030,
                                by = 10)) +
  scale_y_continuous(name = "World Sales (in millions USD)")
  
most_sales %>% 
  head(10) %>% 
  kable(caption = "Fig. 5.2 A table showing the 10 top films of their year.")
```
Note that 9 out of 10 of the films in Fig. 5.2 were released after 2009. Not only that, 6 were distributed by Walt Disney Studios. Twentieth Century Fox is now owned by Disney and will release sequels for the top-spot film "Avatar." While the film 'Spider-Man: No Way Home' was distributed by Sony, they used Marvel Studios, also owned by Disney.

## Sales for Covid (2020-2022)

```{r annual_sale, echo = FALSE, warning = FALSE, fig.cap="Fig. 6.1 Due to COVID-19, movie sales came to a halt in March 2020 as many countries went into lockdown."}
annual_sale <- movies %>% 
  filter(release_year %in% c(2020,2021,2022))
ggplot(annual_sale, 
       aes(x = as.Date(release_date, format="%B %d, %Y"), 
           y = world_sales_mil, 
           color = runtime)) + 
  geom_point() + 
  geom_text_repel(aes(label = title),
                  size = 3, 
                  segment.color = 'grey50', 
                  max.overlaps = 5) + 
  scale_x_date(name = "Date", 
               date_breaks = "3 months", 
               date_labels = "%b %Y") + 
  scale_y_continuous(name = "World Sales (in millions USD)")
```

## Sales in `r params$year`

```{r param_sale, echo = FALSE, warning = FALSE, fig.cap = paste("Fig. 7.1 A view of high-grossing movies in", params$year)}
param_sale <- movies %>% 
  filter(release_year == params$year)
ggplot(param_sale, 
       aes(x = as.Date(release_date, 
                       format = "%B %d, %Y"), 
           y = world_sales_mil)) + 
  geom_point() + 
  labs(x = "Date", y = "World sales (in millions USD)") + 
  geom_text_repel(data = subset(param_sale, world_sales_mil > 1000),
            aes(label = title),
            size = 3,
            segment.color = 'grey50',
            max.overlaps = 10) +
  scale_x_date(expand = expansion(mult = 0.15), 
               date_breaks = "3 months", 
               date_labels = "%b %Y") +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.1)),
                     breaks = seq(from = 0, 
                                  to = (max(param_sale$world_sales_mil) + 100), 
                                  by = 200))
param_sale %>% 
  filter(world_sales_mil > 500) %>% 
  select(title, release_date, world_sales_mil, license, runtime) %>% 
  arrange(world_sales_mil) %>% 
  kable(caption = paste("Fig. 7.2 Movies in", params$year, "with world sales above 500 million USD"))
```

## Sales by Distributor
```{r distributor_sales, echo = FALSE, message = FALSE, fig.width = 8, fig.height = 8, fig.cap = "Fig. 8.1 Total world sales per year by the top 5 distributors for 2000 - present"}
top_5 <- movies %>%
  group_by(distributor) %>%
  summarize(total_sales = sum(world_sales_mil)) %>%
  arrange(desc(total_sales)) %>%
  head(5) %>%
  pull(distributor)
distributor_sales <- movies %>%
  filter(distributor %in% top_5 & release_year > 2000) %>%
  group_by(distributor, release_year) %>%
  summarize(total_sales = sum(world_sales_mil))
ggplot(distributor_sales,
       aes(x = release_year,
           y = total_sales,
           group = distributor,
           color = distributor)) +
  geom_line() +
  labs(x = "Year",
       y = "Total World Sales (in millions USD per year") +
  guides(color = "none") +
  scale_x_discrete(expand = expansion(mult = 0.1),
                   breaks = seq(from = 2000,
                                to = 2030,
                                by = 5)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)),
                     breaks = seq(from = 0, 
                                  to = (max(distributor_sales$total_sales) + 100), 
                                  by = 500)) +
  facet_wrap(~distributor,
             nrow = 2)
```